Tweet Classification: Filtering \\ of Twitter for Company-relevant Tweets

### 2012-12-07 paper/paper.tex
paper

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{@Apple} please make a squid \textbf{\#emoji} so I can put \\
one next to \textbf{@squidneyy22}'s name\\ \hline
Maybe \textbf{\#Apple} will release \textbf{\#iTunes11} \\
tomorrow. Maybe? \\ \hline
\end{tabular}
\caption{Hashtags and username references (note that Apple Inc. does not own the handle @Apple, though it is commonly used this way)}
\label{tab:hashtag_username}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{Nvm} i'll buy you an ipad \textbf{\^{}-\^{}} \\ \hline
\#Apple fires `maps' project manager \textbf{loooool} big \\
 news story here in San Francisco! Go figure. \textbf{:-)} \\ \hline
Think I might have a \textbf{gf}! \textbf{O\_o} oh \\
lord\textbf{..} Lets see how this goes. \\ \hline
\end{tabular}
\caption{Examples of emoticons and common abbreviations found on Twitter}
\label{tab:emoticons}
\end{table}

@@

\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
	nice fish and chip.. good pie apple :-) \\ http://t.co/KVDN5r5l \\ \hline
	20 chicken nuggets, chips, big mac, and a \\ vanilla milkshake LIFE IS GOOD \\ \hline
	1 more day whoop plus safari ride yay \\ \hline
	Ashley staples an apple.  `AHHHH!!!! \\ Apple juice in my eye!' \\ \hline
	I ain't seen uncle Mac in a while \\
	\hline
\end{tabular}
\caption{Examples of false positive tweets}
\label{tab:myfirsttable}
\end{table}

@@

After tokenizing these special features in order (URLs, emoticons, usernames, and hashtags) we are left with a fragmented tweet containing myriad punctuation, capitalization, and creative spellings. As done by \citet{potts2011} we normalize the length of all letter sequences greater than two, as in English these are invariably due to user-added emphasis, and condensing the larger set of tokens focuses the feature set. In the following tweet, for instance, `loooool' is normalized to `loool'.
\begin{quote}
\#news can't wait for my signed \#OoRITE2OutNow it's emotional here right now \textbf{loooool} \#OoRITE2OutNow https://t.co/r4n4cn6C 
\end{quote}

@@

Our desired result was a high precision classifier that could be useful for gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a baseline strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement of approximately 9\% in the worst case. The more relevant and interesting result is the 43\% improvement in precision from the baseline seen with our custom Twitter-aware tokenizer and all other Pipes like lowercasing, Stemming and Bigrams trained on the maximum of 1500 instances. This relatively large improvement indicates not only if the tokenization strategy used important, but also that we were relatively successful in creating an effective tokenizer that can be further augmented by standard natural language processing techniques like stemming, removing stop words, applying n-grams, or simply lowercasing the text. We can see in Table \ref{tab:top_pipelines} below that all of top five pipeline combinations include our Twitter-aware tokenizer. In fact, one must go down to the 24th and 33rd ranked combinations to find ones that include the Number-Digit and Whitespace based tokenizers, respectively.

\begin{table}[h]
\centering
\small
\begin{tabular}{|c|l|r|}
	\hline
	\textit{Rank} & \textit{Pipeline Description} & \textit{Precision} \\
	& & \textit{for Apple} \\ \hline
	1 & Twitter Tokenizer, Lowercased, & \\ & Stemmed, Stopwords Removed, Bigrams & 92.2\% \\ \hline
	2 & Twitter Tokenizer, Lowercased, & \\ & Stopwords Removed, Bigrams & 91.6\% \\ \hline
	3 & Twitter Tokenizer, Stopwords Removed, & \\ & Stemmed, Bigrams & 91.6\% \\ \hline
	4 & Twitter Tokenizer, Stemmed, Bigrams & 91.4\% \\ \hline
	5 & Twitter Tokenizer, Stopwords Removed, & \\ & Bigrams & 91.3\% \\ \hline
	\hline
	24 & Alphanumeric Tokenizer, & \\ & Remove Stopwords, Bigrams & 88.2\% \\ \hline
	\hline
	33 & Whitespace Tokenizer, Stemmed, & \\ & Remove Stopwords, Bigrams & 87.7\% \\ \hline
\end{tabular}
\caption{Top performing pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

\subsection{Pipe Effectiveness}

\textit{Stemming}: \\
~\\
\textit{Stopword removal}: \\
~\\
\textit{Bigrams}: \\
~\\
\textit{URL Replacement}: \\
~\\
\textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

@@

\begin{figure}[h]
\centering
\includegraphics[]{img/tokenizers}
\caption{Precision of main tokenizers}
\label{fig:tokenizer_precision}
\end{figure}

We found that our Twitter-aware tokenizer outperformed whitespace tokenization by [THIS AMOUNT] on average across the . Applying stemming and stop word removal to both pipelines increased the accuracy for almost every case, though widening the gap. \\

>>>>

\begin{table}[ht]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{@Apple} please make a squid \textbf{\#emoji} so I can put \\
one next to \textbf{@squidneyy22}'s name\\ \hline
Maybe \textbf{\#Apple} will release \textbf{\#iTunes11} \\
tomorrow. Maybe? \\ \hline
\end{tabular}
\caption{Hashtags and username references (note that Apple Inc. does not own the handle @Apple, though it is commonly used this way)}
\label{tab:hashtag_username}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{|l|}
	\hline
\textbf{Nvm} i'll buy you an ipad \textbf{\^{}-\^{}} \\ \hline
\#Apple fires `maps' project manager \textbf{loooool} big \\
 news story here in San Francisco! Go figure. \textbf{:-)} \\ \hline
Think I might have a \textbf{gf}! \textbf{O\_o} oh \\
lord\textbf{..} Lets see how this goes. \\ \hline
\end{tabular}
\caption{Examples of emoticons and common abbreviations found on Twitter}
\label{tab:emoticons}
\end{table}

@@

\begin{table}[ht]
\centering
\begin{tabular}{|l|}
	\hline
	nice fish and chip.. good pie apple :-) \\ http://t.co/KVDN5r5l \\ \hline
	20 chicken nuggets, chips, big mac, and a \\ vanilla milkshake LIFE IS GOOD \\ \hline
	1 more day whoop plus safari ride yay \\ \hline
	Ashley staples an apple.  `AHHHH!!!! \\ Apple juice in my eye!' \\ \hline
	I ain't seen uncle Mac in a while \\
	\hline
\end{tabular}
\caption{Examples of false positive tweets}
\label{tab:myfirsttable}
\end{table}

@@

After tokenizing these special features in order (URLs, emoticons, usernames, and hashtags) we are left with a fragmented tweet containing myriad punctuation, capitalization, and creative spellings. As done by \citet{potts2011} we normalize the length of all letter sequences greater than two, as in English these are invariably due to user-added emphasis, and condensing the larger set of tokens focuses the feature set. In the following tweet, for instance, `loooool' would be normalized to `loool'.
\begin{table}[h]
\centering
\begin{tabular}{|l|}
	\hline
	\#news can't wait for my signed \#OoRITE2OutNow it's \\ emotional here right now \textbf{loooool} \\ \#OoRITE2OutNow https://t.co/r4n4cn6C \\
	\hline
\end{tabular}
\caption{Example of normalizing repeated characters}
\label{tab:normalize_characters_example}
\end{table}

@@

Our desired result was a high precision classifier that could be useful for gathering a large number of tweets mentioning a company or its products. Looking at an overview of the results in Figure~\ref{fig:chart_all}, we can see all pipeline combinations had over 70\% of ``apple'' labelings correct with only 50 training tweet instances. However, given that a simple strategy of labeling every tweet as ``apple'' would be correct 64\% of the time, this is only an improvement in precision of approximately 9\%. The more relevant and interesting result is the 43\% improvement in precision from the simple strategy seen with our custom Twitter-aware tokenizer and all other Pipes like lowercasing, Stemming and Bigrams trained on the maximum of 1500 instances. Futhermore, we observed 12\% and 13\% increases in precision versus plain Whitespace and Alphanumeric based tokenizers respectively. This relatively large improvement indicates not only if the tokenization strategy used important, but also that we were relatively successful in creating an effective tokenizer that can be further augmented by standard natural language processing techniques like stemming, removing stop words, applying n-grams, or simply lowercasing the text. We can see in Table \ref{tab:top_pipelines} below that all of top five pipeline combinations include our Twitter-aware tokenizer. In fact, one must go down to the 24th and 33rd ranked combinations to find ones that utilize the Alphanumeric and Whitespace based tokenizers, respectively.

\begin{table}[ht]
\centering
\small
\begin{tabular}{|c|l|r|}
	\hline
	\textit{Rank} & \textit{Pipeline Description} & \textit{Precision} \\
	& & \textit{for Apple} \\ \hline
	1 & Twitter Tokenizer, Lowercased, & \\ & Stemmed, Stopwords Removed, Bigrams & 92.2\% \\ \hline
	2 & Twitter Tokenizer, Lowercased, & \\ & Stopwords Removed, Bigrams & 91.6\% \\ \hline
	3 & Twitter Tokenizer, Stopwords Removed, & \\ & Stemmed, Bigrams & 91.6\% \\ \hline
	4 & Twitter Tokenizer, Stemmed, Bigrams & 91.4\% \\ \hline
	5 & Twitter Tokenizer, Stopwords Removed, & \\ & Bigrams & 91.3\% \\ \hline
	\hline
	24 & Alphanumeric Tokenizer, & \\ & Remove Stopwords, Bigrams & 88.2\% \\ \hline
	\hline
	33 & Whitespace Tokenizer, Stemmed, & \\ & Remove Stopwords, Bigrams & 87.7\% \\ \hline
\end{tabular}
\caption{Top performing pipeline combinations for 1500 training instances}
\label{tab:top_pipelines}
\end{table}

% \subsection{Pipe Effectiveness}

% \textit{Stemming}: \\
% ~\\
% \textit{Stopword removal}: \\
% ~\\
% \textit{Bigrams}: \\
% ~\\
% \textit{URL Replacement}: \\
% ~\\
% \textit{Spell checking}: Correcting misspelled words provided no statistically significant change in our results. \\

@@

\begin{figure}[ht]
\centering
\includegraphics[]{img/tokenizers}
\caption{Precision of main tokenizers}
\label{fig:tokenizer_precision}
\end{figure}

We found that our Twitter-aware tokenizer with the full suite of stemming, stopword removal and bigramming outperformed the whitespace tokenization with comparable stemming, stopword removal and bigramming by 5.2\% in the precision metric when trained on 1500 instances. A similar comparison applies to the alphanumeric tokenizer, our Twitter-aware tokenizer was 4.5\% more percise labeling tweets as mentioning Apple. Applying stemming and stop word removal to both pipelines increased each metric slightly, giving rise to the well distributed set of results seen in Figure \ref{fig:chart_all}. Bigramming and other pipes had some effect, but compared to the multiple percentage point increases from changing the tokenization method, they had relatively minor effects individually.
